Exploration

### 2019-01-03 background/exploration.Rmd

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
x_max <- 100
y_max <- 100

lc.df <- read_csv("data/landcover_5km.csv") %>% 
  filter(!is.na(bio1_mean)) %>%
  mutate(x=as.integer(factor(.$left)),
         y=as.integer(factor(.$top,
                             levels=rev(levels(factor(.$top))))),
         x_y=paste(x, y, sep="_"))
env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
                                 y=1:max(lc.df$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
         temp2=temp^2,
         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
         prec2=prec^2,
         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
         inbd=!is.na(match(.$x_y, lc.df$x_y)),
         id=row_number(), 
         id.inbd=min_rank(na_if(inbd*id, 0)),
         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
env.rct[is.na(env.rct)] <- 0
env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
rm(env.rct)
```

@@

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
n0 <- 100
regPlots <- T
ipmPlots <- T
nPlots <- 5
z.rng <- c(1, 12)
n <- 20  # ncells in IPM matrix
tmax <- 10
cells <- n_distinct(lc.df$CellID)
n.cell <- length(cells)
sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
                    s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(2, -.1, 2, -.1, 2, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(-2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, -1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.03, # p(establishment)
                    NDD=T, # negative density dependence: p_est=p_est/N
                    NDD_n=n0/2, # mean number of recruits if NDD
                    rcr_SB=0.5, # p(recruit from seedbank)
                    rcr_dir=0.5, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    sdd_max=5, # max SDD distance in cells
                    sdd_rate=1 # SDD dispersal rate
                    ))
sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
                        g.p=list(sdd.max=sp_i$p$sdd_max, 
                                 sdd.rate=sp_i$p$sdd_rate, 
                                 bird.hab=rep(1,5)))
```

@@

sample.cells <- sample((1:n.cell)[lam.df$Surv.S>0], nPlots)
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(U$lo, U$hi, length.out=100), size2=size^2, size3=size^3)
  E.t <- xx %>% add_column(yr=NA, s=NA, g=NA, fl=NA, seed=NA)
  xxx <- seq(0, U$hi*.5, length.out=100)
  for(i in sample.cells) {
    z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
    pt.col <- rgb(0,0,0,1/(nrow(z.t)^.25))
    E.t <- sim_expected(1, xx$size, 1:100, E.t, p, map(X, ~.[i,]), n_z)
    main.i <- paste("Cell", i)
    w1 <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=pt.col, 
         xlim=c(U$lo, U$hi),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      lines(E.t$size, E.t$s, col=2)
    plot(sizeNext~size, data=z.t, pch=20, col=pt.col, ylim=c(U$lo,U$hi), 
         xlim=c(U$lo, U$hi),
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      lines(E.t$size, E.t$g, col=2)
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col, xlim=c(U$lo, U$hi),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      lines(E.t$size, E.t$fl, col=2)
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, xlim=c(U$lo, U$hi),
         ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      lines(E.t$size, E.t$seed, col=2)
    plot(U$y, stable.dist1, ylab="", type="l", col=2, 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    hist(with(S$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red")
  }
}
if(ipmPlots) {
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  x <- c(0, U$y)
  for(i in sample.cells) {
   z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
   pt.col <- rgb(1,1,1,1/(nrow(z.t)^.25))
   main.i <- paste("Cell", i)
   z.t$y <- U$y[findInterval(z.t$size, U$y, all.inside=T)]
   totSd <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
   image.plot(x, x, t(U$Ps[,,i]), main=paste("P matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, col=pt.col, pch=20)
    abline(a=0, b=1, col="white", lty=2)
   image.plot(x, x, t(U$Fs[,,i]), main=paste("F matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(totSd$y, totSd$seed/max(totSd$seed)*U$hi*.9, col="white", pch=20)
   image.plot(x, x, t(sens(U$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
   image.plot(x, x, t(elas(U$IPMs[,,i])), main=paste("Elasticity:", main.i),
              xlab="Size (t)", ylab="Size (t+1)")
   points(sizeNext~size, data=z.t, col=pt.col, pch=20)
  }
}
par(mfrow=c(2,2), mar=c(4,4,2,1))
x <- c(0, U$y)
S_d <- map_dfr(S$d, data.frame) %>% sample_n(1000)
S_d$y <- U$y[findInterval(S_d$size, U$y, all.inside=T)]
totSd <- group_by(S_d, y) %>% summarise(seed=sum(seed, na.rm=T))
pt.col <- rgb(1,1,1,1/(nrow(S_d)^.25))
image.plot(x, x, t(apply(U$Ps, 1:2, mean)), main="P matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
points(sizeNext~size, data=S_d, col=pt.col, pch=20)
image.plot(x, x, t(apply(U$Fs, 1:2, mean)), main="F matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
points(totSd$y, totSd$seed/max(totSd$seed)*U$hi*.9, col="white", pch=20)
image.plot(x, x, sens(t(apply(U$IPMs, 1:2, mean))), main="Sensitivity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, elas(t(apply(U$IPMs, 1:2, mean))), main="Elasticity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
points(sizeNext~size, data=S_d, col=pt.col, pch=20)

```

@@

## Observed datasets
The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data. The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution. The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
```{r O_setup}
O_n <- list(Mx=500, CA=20, IPM=20) # number of cells in sample
O_yr <- list(Mx=tmax, CA=(tmax-10):tmax, IPM=tmax) # years to sample from
P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
n_sim <- 15
```

### Correlative: MaxEnt
```{r O_Mx}
library(dismo)
Mx.P <- sample(P.i, O_n$Mx, replace=F)
pres <- 1:nrow(lam.df) %in% Mx.P
Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
                           (MxPred==0 & lambda < 1))
ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
```
### Mechanistic: CA
```{r O_CA}

```
### Mechanistic: IPM
```{r O_IPM}
IPM.P <- sample(P.i, O_n$IPM, replace=FALSE)
IPM.d <- vector("list", O_n$IPM)
# sample individuals from each cell rather than full census -- proportion?
for(i in IPM.P) {
  IPM.d[[i]] <- data.frame(S$d[[i]]) %>% filter(yr %in% O_yr$IPM) %>%
    add_column(id.inbd=i) %>%
    full_join(env.df[i,], by="id.inbd")
}
IPM.d <- do.call(rbind, IPM.d)
p.ipm <- p
```

@@

## Fitting the models
Each observed dataset *O* is used to predict the distribution of species *S*. The predicted distribution *P* is compared to the known truth *S* to assess the predictive accuracy of each modeling approach, given any limitations, uncertainty, or bias imposed in *O*. 
### Correlative: MaxEnt
```{r P_Mx}

```
### Mechanistic: CA
```{r P_CA}

@@

IPM.fit.sum <- lam.df %>% 
  summarise(mnLamDiff=mean(lambda.f-lambda),
            mdLamDiff=median(lambda.f-lambda),
            sdLamDiff=sd(lambda.f-lambda),
            pctLamCorr=mean((lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1)),
            mnLamSDiff=mean(lam.S.f-lam.S, na.rm=T),
            mdLamSDiff=median(lam.S.f-lam.S, na.rm=T),
            sdLamSDiff=sd(lam.S.f-lam.S, na.rm=T),
            pctLamSCorr=mean((lam.S>1 & lam.S.f>1) | (lam.S<1 & lam.S.f<1)),
            mnNDiff=mean(N.S.f-N.S),
            mdNDiff=median(N.S.f-N.S),
            sdNDiff=sd(N.S.f-N.S),
            pctNCorr=mean((N.S>0 & N.S.f>0) | (N.S==0 & N.S.f==0)),
            mnSurvDiff=mean(Surv.S.f-Surv.S),
            mdSurvDiff=median(Surv.S.f-Surv.S),
            sdSurvDiff=sd(Surv.S.f-Surv.S),
            pctSurvCorr=mean((Surv.S>0 & Surv.S.f>0) | (Surv.S==0 & Surv.S.f==0)),
            mnRcrDiff=mean(Rcr.S.f-Rcr.S),
            mdRcrDiff=median(Rcr.S.f-Rcr.S),
            sdRcrDiff=sd(Rcr.S.f-Rcr.S),
            pctRcrCorr=mean((Rcr.S>0 & Rcr.S.f>0) | (Rcr.S==0 & Rcr.S.f==0)))
str(IPM.fit.sum)
```

>>>>

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
x_max <- 20
y_max <- 20

lc.df <- read_csv("data/landcover_5km.csv") %>% 
  filter(!is.na(bio1_mean)) %>%
  mutate(x=as.integer(factor(.$left)),
         y=as.integer(factor(.$top,
                             levels=rev(levels(factor(.$top))))),
         x_y=paste(x, y, sep="_"))
env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
                                 y=1:max(lc.df$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
         temp2=temp^2,
         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
         prec2=prec^2,
         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
         inbd=!is.na(match(.$x_y, lc.df$x_y)),
         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
         lon=lc.df$left[match(.$x_y, lc.df$x_y)]) %>%
  filter(y >= (max(.$y) - y_max) & x <= x_max) %>%
  mutate(id=row_number(), 
         id.inbd=min_rank(na_if(inbd*id, 0)))
env.rct[is.na(env.rct)] <- 0
env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
n.cell <- sum(env.rct$inbd)
```

@@

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
n0 <- 100
regPlots <- T
ipmPlots <- T
nPlots <- 5
z.rng <- c(1, 12)
n <- 20  # ncells in IPM matrix
tmax <- 10
sp_i <- list(p=list(s_z=c(-8, 2.1, -.09), # b1 + b2*z + b3*z^2
                    s_x=c(2, -.1, -2, -.1, 2, -2, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(2, -.1, 2, -.1, 2, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2
                    fl_x=c(-2, -.1, -2, -.1, 1, 1), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, -1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.03, # p(establishment)
                    NDD=T, # negative density dependence: p_est=p_est/N
                    NDD_n=n0/2, # mean number of recruits if NDD
                    rcr_SB=0.5, # p(recruit from seedbank)
                    rcr_dir=0.5, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    sdd_max=5, # max SDD distance in cells
                    sdd_rate=1 # SDD dispersal rate
                    ))
sp_i$p$p_emig <- pexp(0.5, sp_i$p$sdd_rate, lower.tail=F) # p(seed emigrates)
# sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
                        g.p=list(sdd.max=sp_i$p$sdd_max, 
                                 sdd.rate=sp_i$p$sdd_rate, 
                                 bird.hab=rep(1,5)))
```

@@

sample.cells <- sample((1:n.cell)[lam.df$Surv.S>0], nPlots)
if(regPlots) {
  par(mfrow=c(3,2), mar=c(4,4,2,1))
  xx <- tibble(size=seq(U$lo, U$hi, length.out=100), size2=size^2, size3=size^3)
  E.t <- xx %>% add_column(yr=NA, s=NA, g=NA, fl=NA, seed=NA)
  xxx <- seq(0, U$hi*.5, length.out=100)
  for(i in sample.cells) {
    z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
    pt.col <- rgb(0,0,0,1/(nrow(z.t)^.25))
    E.t <- sim_expected(1, xx$size, 1:100, E.t, p, map(X, ~.[i,]), n_z)
    main.i <- paste("Cell", i)
    w1 <- Re(eigen(U$IPMs[,,i])$vectors[,1])[-1]; stable.dist1 <- w1/sum(w1)
    plot(surv~size, data=z.t, ylim=c(0,1), pch=20, col=pt.col, 
         xlim=c(U$lo, U$hi),
         main=paste("Surv:", main.i), xlab="Size (t)", ylab="P(surv to t+1)")
      lines(E.t$size, E.t$s, col=2)
    plot(sizeNext~size, data=z.t, pch=20, col=pt.col, ylim=c(U$lo,U$hi), 
         xlim=c(U$lo, U$hi),
         main=paste("Growth:", main.i), xlab="Size (t)", ylab="Size (t+1)")
      lines(E.t$size, E.t$g, col=2)
      abline(a=0, b=1, lty=2)
    plot(fl~size, data=z.t, pch=20, ylim=c(0,1), col=pt.col, xlim=c(U$lo, U$hi),
         main=paste("P(Flower):", main.i), xlab="Size (t)", ylab="P(flower)")
      lines(E.t$size, E.t$fl, col=2)
    plot(seed~size, data=z.t, pch=23-z.t$fl*3, col=pt.col, xlim=c(U$lo, U$hi),
         ylim=range(c(E.t$seed, z.t$seed), na.rm=TRUE),
         main=paste("nSeeds:", main.i), xlab="Size (t)", ylab="Number of seeds")
      lines(E.t$size, E.t$seed, col=2)
    plot(U$y, stable.dist1, ylab="", type="l", col=2, 
         main=paste("Stable size distribution:", main.i), xlab="Size (t)")
    hist(with(S$d[[i]], sizeNext[is.na(size)]), col="grey", freq=F, 
         main=paste("Recruit sizes:", main.i), xlab="Recruit size (t+1)")
      lines(xxx, dnorm(xxx, p$rcr_z[1], p$rcr_z[2]), col="red")
  }
}
if(ipmPlots) {
  par(mfrow=c(2,2), mar=c(4,4,2,1))
  x <- c(0, U$y)
  for(i in sample.cells) {
   z.t <- data.frame(S$d[[i]]) %>% filter(!is.na(size) & yr==tmax)
   pt.col <- rgb(1,1,1,1/(nrow(z.t)^.25))
   main.i <- paste("Cell", i)
   z.t$y <- U$y[findInterval(z.t$size, U$y, all.inside=T)]
   totSd <- group_by(z.t, y) %>% summarise(seed=sum(seed, na.rm=T))
   image.plot(x, x, t(U$Ps[,,i]), main=paste("P matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(sizeNext~size, data=z.t, col=pt.col, pch=20)
    abline(a=0, b=1, col="white", lty=2)
   image.plot(x, x, t(U$Fs[,,i]), main=paste("F matrix:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
    points(totSd$y, totSd$seed/max(totSd$seed)*U$hi*.9, col="white", pch=20)
   image.plot(x, x, t(sens(U$IPMs[,,i])), main=paste("Sensitivity:", main.i), 
              xlab="Size (t)", ylab="Size (t+1)")
   image.plot(x, x, t(elas(U$IPMs[,,i])), main=paste("Elasticity:", main.i),
              xlab="Size (t)", ylab="Size (t+1)")
   points(sizeNext~size, data=z.t, col=pt.col, pch=20)
  }
}
par(mfrow=c(2,2), mar=c(4,4,2,1))
x <- c(0, U$y)
S_d <- map_dfr(S$d, data.frame) %>% sample_n(1000)
S_d$y <- U$y[findInterval(S_d$size, U$y, all.inside=T)]
totSd <- group_by(S_d, y) %>% summarise(seed=sum(seed, na.rm=T))
pt.col <- rgb(1,1,1,1/(nrow(S_d)^.25))
image.plot(x, x, t(apply(U$Ps, 1:2, mean)), main="P matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
points(sizeNext~size, data=S_d, col=pt.col, pch=20)
image.plot(x, x, t(apply(U$Fs, 1:2, mean)), main="F matrix: mean",
           xlab="Size (t)", ylab="Size (t+1)")
points(totSd$y, totSd$seed/max(totSd$seed)*U$hi*.9, col="white", pch=20)
image.plot(x, x, sens(t(apply(U$IPMs, 1:2, mean))), main="Sensitivity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
image.plot(x, x, elas(t(apply(U$IPMs, 1:2, mean))), main="Elasticity: mean",
           xlab="Size (t)", ylab="Size (t+1)")
points(sizeNext~size, data=S_d, col=pt.col, pch=20)
```

@@

# Exploration of options for the sampling component

```{r loadSpecies}
env.in <- readRDS("out/sp1_env_in.rds")
n.cell <- nrow(env.in)
p <- readRDS("out/sp1_p.rds")
sdd.pr <- readRDS("out/sp1_sdd.rds")
U <- readRDS("out/sp1_U.rds")
S <- readRDS("out/sp1_S.rds")
lam.df <- readRDS("out/sp1_lam_df.rds")
```

## Observed datasets
The simulated species *S* is sampled appropriately in order to fit each SDM model. The observed datasets,*O*, may be random or intentionally biased to mimic common issues with empirical data. The realized simulated data generated above (`d` and its abundance summary `N_sim`) are the underlying true distribution. The samples will be drawn from occupied cells in the final time steps where the species is approximately at equilibrium.
```{r O_setup}
O_n <- list(Corr=100, Mech=20) # number of cells in sample
O_yr <- list(Mx=p$tmax, CA=(p$tmax-10):p$tmax, IPM=p$tmax) # years to sample from
P.i <- which(lam.df$Surv.S > 0) # presences: surviving individuals > seedlings
Mech.sample <- sample(P.i, O_n$Mech, replace=FALSE)
Corr.sample <- sample(P.i, O_n$Corr, replace=FALSE)
n_sim <- 15
```

### Correlative: MaxEnt
```{r O_Mx}
pres <- 1:nrow(lam.df) %in% Corr.sample
```
### Mechanistic: CA
```{r O_CA}

```
### Mechanistic: IPM
```{r O_IPM}
IPM.d <- vector("list", O_n$Mech)
# sample individuals from each cell rather than full census -- proportion?
for(i in Mech.sample) {
  IPM.d[[i]] <- data.frame(S$d[[i]]) %>% filter(yr %in% O_yr$IPM) %>%
    add_column(id.inbd=i) %>%
    full_join(env.in[i,], by="id.inbd")
}
IPM.d <- do.call(rbind, IPM.d)
p.ipm <- p
```

@@

# Exploration of options for the fitting component

## Fitting the models
Each observed dataset *O* is used to predict the distribution of species *S*. The predicted distribution *P* is compared to the known truth *S* to assess the predictive accuracy of each modeling approach, given any limitations, uncertainty, or bias imposed in *O*. 
### Correlative: MaxEnt
```{r P_Mx}
Mx.out <- maxent(x=lam.df[,c(1,3,5:9)], p=pres)
lam.df$MxPred <- 1:nrow(lam.df) %in% row.names(Mx.out@presence)
lam.df$MxCorrect <- with(lam.df, (MxPred==1 & lambda > 1) |
                           (MxPred==0 & lambda < 1))
ggplot(lam.df, aes(x=lon, y=lat, fill=MxCorrect)) + geom_tile()
```
### Mechanistic: CA
```{r P_CA}

@@

########
## comparison with truth
########
IPM.fit.sum <- lam.df %>% 
  summarise(mnLamDiff=mean(lambda.f-lambda),
            mdLamDiff=median(lambda.f-lambda),
            sdLamDiff=sd(lambda.f-lambda),
            pctLamCorr=mean((lambda>1 & lambda.f>1) | (lambda<1 & lambda.f<1)),
            mnLamSDiff=mean(lam.S.f-lam.S, na.rm=T),
            mdLamSDiff=median(lam.S.f-lam.S, na.rm=T),
            sdLamSDiff=sd(lam.S.f-lam.S, na.rm=T),
            pctLamSCorr=mean((lam.S>1 & lam.S.f>1) | (lam.S<1 & lam.S.f<1)),
            mnNDiff=mean(N.S.f-N.S),
            mdNDiff=median(N.S.f-N.S),
            sdNDiff=sd(N.S.f-N.S),
            pctNCorr=mean((N.S>0 & N.S.f>0) | (N.S==0 & N.S.f==0)),
            mnSurvDiff=mean(Surv.S.f-Surv.S),
            mdSurvDiff=median(Surv.S.f-Surv.S),
            sdSurvDiff=sd(Surv.S.f-Surv.S),
            pctSurvCorr=mean((Surv.S>0 & Surv.S.f>0) | (Surv.S==0 & Surv.S.f==0)),
            mnRcrDiff=mean(Rcr.S.f-Rcr.S),
            mdRcrDiff=median(Rcr.S.f-Rcr.S),
            sdRcrDiff=sd(Rcr.S.f-Rcr.S),
            pctRcrCorr=mean((Rcr.S>0 & Rcr.S.f>0) | (Rcr.S==0 & Rcr.S.f==0)))
str(IPM.fit.sum)
```
